---
title: "Sta303 - TidyTuesday 1"
author: "Erdong Zhang"
date: "2021/3/4"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("tidytuesdayR")
library(tidyverse)
library(scales)
tuesdata <- tidytuesdayR::tt_load('2021-02-23')

```

```{r}
earn<- tuesdata$earn
employed <- tuesdata$employed

employed %>%
  count(race_gender, sort = TRUE)
# The original data does not categorize gender and race
# We need to specify the categories

employed<- tuesdata$employed %>%
  mutate(dimension = case_when(
    race_gender == "TOTAL" ~ "Total",
    race_gender %in% c("Men", "Women") ~ "Gender",
    TRUE ~ "Race"
  ))
```

Bar plot that shows the composition of number of employed in major industries.
```{r}
employed_clean<- employed %>%
  filter(!is.na(employ_n)) %>%
  mutate(industry = fct_lump(industry, 15, w = employ_n), # The top 8 industries with largest workers.
         industry = fct_reorder(industry, employ_n, sum)) # Aggregate the workers in each industries.
# Bar plot
employed_clean %>%
  filter(dimension == "Total") %>%
  ggplot(aes(x = year, y = employ_n, fill = industry)) +
  geom_col() +
  scale_y_continuous(labels = comma) +
  labs(y = "Number employed in industry",
       x = "Year",
       title = "The composition of number of employed in major industries")
```

Facet bar plots that show the total number of workers in each industry for each year.
```{r}
# Facet bar plots that show the total number of workers in each industry for each year.
employed_clean %>%
  filter(dimension == "Total") %>%
  ggplot(aes(x = year, y = employ_n, fill = industry)) +
  geom_col() +
  facet_wrap(~industry, scales = "free_y") +
  theme(legend.position =  "none") +
  scale_y_continuous(labels = comma) +
  labs(y = "Number employed in industry",
       x = "Year") 
```

```{r}
employed_gender <- employed_clean %>%
  filter(dimension == "Gender") %>%
  filter(!is.na(employ_n)) %>%
  group_by(industry, year, race_gender)%>%
  summarise(employ_n = sum(employ_n), .groups = "drop")

employed_gender %>%  
  ggplot(aes(x = year, y = employ_n, color = race_gender)) +
  geom_line() +
  facet_wrap(~industry, scales = "free_y") +
  scale_y_continuous(labels = comma) +
  expand_limits(y = 0)
  labs(y = "Number employed in industry",
       x = "Year")
```

Since there is not so much variation in employment between male and female from 2015 to 2019. Let's look at 2019 to 2020 specifically.
```{r}
industry_2019_2020 <- employed_clean %>%
  filter(year %in% c(2019, 2020)) %>%
  group_by(year, industry, dimension, race_gender) %>%
  summarise(employ_n = sum(employ_n), .groups = "drop")
  
```

A line plot that shows the number of employment from 2019 to 2020 in each industry.
```{r}
industry_2019_2020 %>%
  filter(dimension == "Total") %>%
  ggplot(aes(x = year, y = employ_n, color = industry)) +
  geom_line() +
  expand_limits(y = 0) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = 2019:2020) +
  scale_color_discrete(guide = guide_legend(reverse = TRUE)) +
  labs(x = "year",
       y = "Number of employed")
```

Estimate the effect of pandemic on employment for each industry
```{r}
compare_2019_2020 <- industry_2019_2020 %>%
  arrange(year) %>%
  group_by(industry, dimension, race_gender) %>%
  summarize(ratio = last(employ_n)/ first(employ_n),
            change = ratio - 1,
            employed_2019 = first(employ_n), .groups = "drop") %>%
  ungroup()

compare_2019_2020 %>%
  filter(dimension == "Total") %>%
  mutate(industry = fct_reorder(industry,change)) %>%
  ggplot(aes(x = change, y = industry)) +
  geom_col() +
  scale_x_continuous(labels = percent) +
  labs(title = "Industries that suffered in employment drop in 2020",
       x = "Shift in number employed from 2019 to 2020",
       y = "")

```

Estimate the effect of pandemic on employment for each industry on gender.
```{r}
compare_2019_2020 %>%
  filter(dimension == "Gender") %>%
  mutate(industry = fct_reorder(industry,change)) %>%
  ggplot(aes(x = change, y = industry, fill = race_gender)) +
  geom_col(position = "dodge") +
  scale_x_continuous(labels = percent) +
  labs(title = "Industries that suffered in employment drop in 2020",
       x = "Shift in number employed from 2019 to 2020 by gender",
       y = "",
       fill = "Gender")
```

Estimate the effect of pandemic on employment for each industry on different races.
```{r}
compare_2019_2020 %>%
  filter(dimension == "Race") %>%
  mutate(industry = fct_reorder(industry,change)) %>%
  ggplot(aes(x = change, y = industry, fill = race_gender)) +
  geom_col(position = "dodge") +
  scale_x_continuous(labels = percent) +
  labs(title = "Industries that suffered in employment drop in 2020",
       x = "Shift in number employed from 2019 to 2020 by gender",
       y = "",
       fill = "Race")
```










